---
title: "LBW tutorial"
author: "Joe Brew"
date: "10/01/2014"
output: html_document
---

For this activity, we'll be examining the incidence of low birth weight around the state of Florida.  Data are from Florida Charts (<http://www.floridacharts.com>), a site well worth checking out for lots of aggregated health data in our state.

## 1. Get the data.

Create a folder named lbw_tutorial and download the data into that folder.  The specific dataset for this activity is at <https://github.com/joebrew/misc/blob/master/lbw_tutorial/lbw.csv>.  If you're familiar with github, you can clone the joebrew/misc directory entirely; if not, simply go to the previous link and download the data as a csv file (right click on the "Raw" button and save as a file). 

## 2. Start an R script.

In the lbw_tutorial directory, create an R script named lbw.R.

## 3. Read in the data. 

You'll do this in two steps.  First use 
```{r, eval = FALSE}
setwd()
```
to set your working directory, then use
```{r, eval = FALSE}
read.csv()
```
to read in the data.  Assign it to an object named lbw.

```{r, echo = FALSE, eval = TRUE, message = FALSE}
setwd('~/Documents/misc/lbw_tutorial')
lbw <- read.csv("lbw.csv")
```

## 4. Examine the data's structure.

Ensure that numerical / character / categorical variables are as they should be.  Do this using 
```{r, eval = FALSE} 
str() 
head()
tail()
summary()
```

If not, transform them using
```{r, eval = FALSE}
as.character()
as.numeric()
as.factor()
````

## 5. Explore the data's contents

First, use dplyr to arrange (sort) the data by percent poverty
```{r results='hide', message = FALSE}
library(dplyr)
lbw <- 
  lbw %>%
  arrange(percent_poverty)
```

Make an ordered barplot of the percent_poverty variable using 
```{r, eval = FALSE}
barplot()
```
Use the names.arg, cex.names and las = 3 arguments in the barplot() function to print the county names vertically below the chart. If you want a challenge, color Alachua's bar red and print a horizontal line at Alachua's poverty rate.

```{r, eval = TRUE, echo = FALSE}
#### !!!!!!!!!!
mycols <- rep(adjustcolor("darkgreen", alpha.f = 0.1), nrow(lbw))
mycols[which(lbw$county == "Alachua")] <- adjustcolor("darkred", alpha.f =0.5)
barplot(lbw$percent_poverty,
        col = mycols,
        border = adjustcolor("black", alpha.f = 0.1),
        names.arg = lbw$county,
        las = 3,
        cex.names = 0.5)
abline(h = lbw$percent_poverty[which(lbw$county == "Alachua")],
       col = adjustcolor("darkred", alpha.f = 0.5))
```

Make a histogram of percent_poverty.

```{r, eval = TRUE, echo = FALSE}
hist(lbw$percent_poverty, col = adjustcolor(rainbow(20), alpha.f = 0.4),
     border = "grey",
     xlab = "Poverty rate", breaks = 20,
     main = "Distribution of poverty rate in Florida by county")
```

Re-arrange our data by population
```{r, eval = TRUE, echo = TRUE}
lbw <- lbw %>%
  arrange(total_pop)
```


## 6. Create some useful new variables
Create variables named white_per, black_per and other_per. Also create lbw_per. These should be percent variables and should be easy to construct using simple divison and the total_pop vector and live_births vector.  


```{r, eval = TRUE, echo = FALSE}
#### !!!!!!!!!!
lbw <- 
  lbw %>%
  mutate(white_per = white / total_pop * 100,
         black_per = black / total_pop * 100,
         other_per = other / total_pop * 100,
         lbw_per = lbw_births / live_births * 100)
```



## 7. Plot the relationship between the poverty rate and lbw rate.  Use

```{r, eval = FALSE}
plot(x, y)
```

If you want to get fancy, make the size (cex) a function of a county's population size, as in the below plot

```{r, eval = TRUE, echo = FALSE}
#### !!!!!!!!!!
plot(lbw$percent_poverty, lbw$lbw_per,
     xlab = "Percent poverty",
     ylab = "LBW rate",
     pch = 16, 
     col = adjustcolor("red", alpha.f = 0.3),
     cex = lbw$total_pop ^ (1/4) / 10)
points(lbw$percent_poverty, lbw$lbw_per,
     col = adjustcolor("black", alpha.f = 0.5),
     cex = lbw$total_pop ^ (1/4) / 10)
```


## 8. Plot the relationship between the percent of residents who are white and the lbw rate.

```{r, eval = TRUE, echo = FALSE}
#### !!!!!!!!!!
plot(lbw$white_per, lbw$lbw_per,
     xlab = "Percent white",
     ylab = "LBW rate",
     pch = 16, 
     col = adjustcolor("darkblue", alpha.f = 0.3),
     cex = lbw$total_pop ^ (1/4) / 10)
points(lbw$white_per, lbw$lbw_per,
     col = adjustcolor("black", alpha.f = 0.5),
     cex = lbw$total_pop ^ (1/4) / 10)
```

##9. Construct a linear model to describe the relationship between lbw rate (y) and two independent variables: percent of residents who are black and the percent of residents who are poor. Name your linear model "fit", and then summarize, plot it and calculate confidence intervals

```{r, eval = FALSE}
fit <- lm()
summary(fit)
plot(fit)
confint(fit)
```

```{r, eval = TRUE, echo = FALSE}
#### !!!!!!!!!!
fit <- lm(lbw_per ~ black_per + percent_poverty, data = lbw)
```

## 10. Intepret the regression.  At the county-level, after adjustment, what factors are significantly associated with lbw rate?
